---
title: "Data Wrangling"
author: "Danielle Lawrence"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Chapter 9  
Wrangling includes: import, tidy, and transform  

Relational data will give you tools for working with multiple interrelated datasets.  

Strings will introduce regular expressions, a powerful tool for manipulating strings.  

Factors are how R stores categorical data. They are used when a variable has a fixed set of possible values, or when you want to use a non-alphabetical ordering of a string.  

Dates and times will give you the key tools for working with dates and date-times.  



### Chapter 10  


```{r, include=FALSE}
library(tidyverse)
```
  
10.2  
 as_tibble(data) to make a data frame into a tibble  
 tibble() creates a new tibble from individual vectors  
 another way is tribble()  
 
10.3.1 Printing  
tibbles print first 10 rows and as many columns that an fit  
can change how much shows when print with print()  

10.3.2 Subsetting  
df <- tibble(
  x = runif(5),
  y = rnorm(5))  
to extract by name: df$x or df[["x"]]  
to extract by position: df[[1]]  
to use these in a pipe you have to add .  

10.5 Exercises  
1. If only the first 10 rows are printed  

2. 
```{r}
df <- as_tibble(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```
it resulted in an error when I try to do a tibble  

3. data$var or data[["var"]]  

4.  
  -- data$1 or data[["1"]]   
  -- x=1 y=2  
  -- mutate the data into 3 = 2/1  
  -- rename them  

5. changes named vectors into names and values  

6. you can use print(n=?) to change what prints out  



Chapter 11  
many ways to read in data, but we focus on read_csv()  
it prints out a column that gives the name and type of each column  
col_names = FALSE if they arent named  
skip = n to skip the first n lines  

11.2.2 Exercises  
1. read_delim()  
2. na and col_names  
3. fwf_widths() or fwf_positions  
4. read_delim()  
5. there are different separations within the same code  


11.3  
parse_*()  

8 important ones:  
parse_logical() and parse_integer() parse logicals and integers respectively. There’s basically nothing that can go wrong with these parsers so I won’t describe them here further.  

parse_double() is a strict numeric parser, and parse_number() is a flexible numeric parser. These are more complicated than you might expect because different parts of the world write numbers in different ways.  

parse_character() seems so simple that it shouldn’t be necessary. But one complication makes it quite important: character encodings.  

parse_factor() create factors, the data structure that R uses to represent categorical variables with fixed and known values.  

parse_datetime(), parse_date(), and parse_time() allow you to parse various date & time specifications. These are the most complicated because there are so many different ways of writing dates.  
  

11.3.5 Exercises  
1. date_names, date_format, time_format  
2. if they're the same, its an error. if decimal is comman, then grouping is period. then it can be flipped  
3. used to describe date and time formats. like data from another country.  
4. locale(date_format = "%d/%m/%Y")  
5. read_csv() uses a comma, read_csv2() uses semicolons  
6. ASCII  
7.  
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"

parse_date(d1, "%B %d, %Y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3, "%d-%b-%Y")
parse_date(d4, "%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1, "%H%M")








